---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  gganimate = list(
    nframes = 50
  ),
  out.width = '100%'
)
```

*gganimate* is an extension of the *grammar of graphics*, as implemented by the
[*ggplot2*](https://ggplot2.tidyverse.org) package, that adds support for 
declaring animations using an API familiar to users of *ggplot2*.

> The following introduction assumes familiarity with *ggplot2* to the extend 
  that constructing static plots and reading standard *ggplot2* code feels
  natural. If you are new to both *ggplot2* and *gganimate* you'll benefit from
  exploring the trove of *ggplot2* documentation, tutorials, and courses 
  avaialble online first (see the 
  [*ggplot2* webpage](https://ggplot2.tidyverse.org/#learning-ggplot2) for some 
  pointers).
  
## Your First Animation
We'll jump right into our first animation. Don't worry too much about 
understanding the code, as we'll dissect it later.

```{r}
library(gganimate)

# We'll start with a static plot
p <- ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point()

plot(p)
```

You go from a static plot made with *ggplot2* to an animated one, simply by 
adding on functions from *gganimate*.

```{r}
anim <- p + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

anim
```

> ❗ `transition_states()` splits up plot data by a discrete varaible and 
  animates between the different states.

As can be seen, very few additions to the plot results in a quite complex 
animation. So what did we do to get this animation? We added a type of 
*transition*. Transitions are functions that interpret the plot data in order to
somehow distribute it over a number of frames. `transition_states()` 
specifically splits the data into subsets based on a variable in the data (here
`Species`), and calculates intermediary data states that ensures a smooth
transition between the states (something refered to as *tweening*). *gganimate*
provides a range of different transitions, but for the rest of the examples
we'll be sticking to `transition_states()` and see how we can modify the output.

## Easing
When `transition_states()` calculates intermediary data for the tweening, it 
needs to decide how the change from one value to another should progress. This
is a concept called *easing*. The default easing is *linear*, but others can be
used, potentially only targeting specific aesthetics. Setting easing is done
with the `ease_aes()` function. The first argument sets the default easing and
subsequent named arguments sets it for specific aesthetics.

```{r}
anim + 
  ease_aes('cubic-in-out') # Slow start and end for a smoother look
```

> ❗ `ease_aes()` defines the  velocity with which aesthetics change during an
  animation.

```{r}
anim + 
  ease_aes('linear', y = 'bounce-out') # Sets special ease for y aesthetic
```

## Labelling
It can be quite hard to understand an animation without any indication as to
what each time point relates to. *gganimate* solves this by providing a set of 
variables for each frame, which can be inserted into plot labels using 
[*glue*](https://glue.tidyverse.org) syntax.

```{r}
anim + 
  ggtitle('Now showing {closest_state}',
          subtitle = 'Frame {frame} of {nframes}')
```

> ❗ Use *glue* syntax to insert frame variables in plot labels and titles.

Different transitions provide different frame variables. `closest_state` only
makes sense for `transition_states()` and is thus only avaialable when that 
transition is used.

## Object Permanence
In the animation above, it appears as if data if a single measurement changes
gradually as the flower being measured on somehow morphs between three different
iris species. This is probably not how Fisher conducted the experiement and got
those numbers. In general, when you make an animation, *graphic elements should
only transition between instances of the same underlying phenomenon*. 
This sounds complicated but it is more or less the same principle that governs 
makes sense to draw a line between two observations. You wouldn't connect 
observations from different iris species, but repeated observations on the same
plant would be ok to connect. Same thing with animations.

Just to make this very clear (it is an important concept). The line plot 
equivalent of our animation above is:

```{r}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_line(aes(group = rep(1:50, 3)), colour = 'grey') + 
  geom_point()
```

Ugh...

So, how do we fix this and tell *gganimate* to not morph observations from 
different species into each others? The key is the *group aesthetic*. You may be
familiar with this aesthetic from plotitng lines and polygons, but in 
*gganimate* it takes a more central place. Data that have the same group 
aesthetic are interpreted as being linked across states. The semantics of the
group aesthetic in *ggplot2* is such that if it is undefined it will get 
calculated based on the interaction of all discrete aesthetics (sans `label`). 
If none exists, such as in our iris animation, all data will get the same group, 
and will thus be matched by *gganimate*. So, there are two ways to fix our plot:

1. Add some aesthetics that distinguish the different species
```{r}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(colour = Species)) + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)
```

2. Set the group directly
```{r}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(group = seq_along(Species))) + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)
```

> ❗ The group aesthetic defines how the data in a layer is matched across the
  animation.

In general *2)* is preferred as it makes the intend explicit. It also makes it 
possible to match data with different discrete aesthetics such as keeping our 
(now obviously faulty) transition while having different colour for the different
species)

```{r}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(colour = Species, group = 1L)) + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)
```

## Enter and Exit
While we may have made our animation more correct by separating the data from 
the different species, we have also made it quite a bit more boring. Now it
simply appears as three static plots shown one at a time, which is hardly an
attention grabber. If only there were a way to animate the appearance and 
disappearance of data...

Enter the `enter` and `exit` functions. These functions are responsible for 
modifying the state of appearing (entering) and disappearing (exiting) data, so
that the animation can tween from and to the new state. Let's spice up our 
animation a bit:

```{r}
anim <- ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(colour = Species), size = 2) + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

anim + 
  enter_fade() + 
  exit_shrink()
```

> ❗ `enter` and `exit` functions are used to modify the aesthetics of appearing
  and disappearing data so that their entrance or exit may be animated.
  
*gganimate* comes with a range of different functions, and using the 
`enter_manual()` and `exit_manual()` functions you can create your own. Enter 
and exit functions are composable though, so you can often come pretty far by
combining preexisting ones

```{r}
anim + 
  enter_fade() + enter_drift(x_mod = -1) + 
  exit_shrink() + exit_drift(x_mod = 5)
```

## Rendering

