---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  gganimate.nframes = 50
)
```

*gganimate* is an extension of the *grammar of graphics*, as implemented by the
[*ggplot2*](https://ggplot2.tidyverse.org) package, that adds support for 
declaring animations using an API familiar to users of *ggplot2*.

> The following introduction assumes familiarity with *ggplot2* to the extend 
  that constructing static plots and reading standard *ggplot2* code feels
  natural. If you are new to both *ggplot2* and *gganimate* you'll benefit from
  exploring the trove of *ggplot2* documentation, tutorials, and courses 
  avaialble online first (see the 
  [*ggplot2* webpage](https://ggplot2.tidyverse.org/#learning-ggplot2) for some 
  pointers).
  
## Your First Animation
We'll jump right into our first animation. Don't worry too much about 
understanding the code, as we'll dissect it later.

```{r}
library(gganimate)

# We'll start with a static plot
p <- ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point()

plot(p)
```

You go from a static plot made with *ggplot2* to an animated one, simply by 
adding on functions from *gganimate*.

```{r}
anim <- p + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

animate(anim)
```

> ❗ `transition_states()` splits up plot data by a discrete varaible and 
  animates between the different states.

As can be seen, very few additions to the plot results in a quite complex 
animation. 
